---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# autoquarto

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/autoquarto)](https://CRAN.R-project.org/package=autoquarto)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![experimental](https://github.com/openpharma/autoquarto/actions/workflows/experimental.yaml/badge.svg)](https://github.com/openpharma/autoquarto/actions/workflows/experimental.yaml)
[![Test Coverage](https://raw.githubusercontent.com/openpharma/autoquarto/_xml_coverage_reports/data/main/badge.svg)](https://github.com/openpharma/autoquarto/blob/_xml_coverage_reports/data/main/coverage.xml)<!-- badges: end -->

The goal of autoquarto is to provide tools to allow the programmatic, incremental definition and publication of Quarto reports. 

`autoquarto` is very much a work in progress, and is being developed in parallel with the [`rbqmR`](https://openpharma.github.io/rbqmR/main/) package.  As `rbqmR` evolves, so will `autoquarto`.


## Installation

You can install the development version of `autoquarto` from [GitHub](https://github.com/) with:


``` r
# install.packages("devtools")
devtools::install_github("openpharma/autoquarto")
```

`autoquarto` is not currently available on CRAN.

## Usage

```{r, echo=FALSE}
library(autoquarto)
```
The fundamental object used to do this is the S4 `QuartoObject` class, two subclasses of which, `QuartoBook` and `QuartoWebsite` are currently defined.  Of these, testing has started only for `QuartoBook`.

The `QuartoObject` class has 4 slots, named `templateSearchPath`, `variables`, `chapters` and `type`.

* The `chapters` slot contains the names of the files that define the chapters, or sections, of the report.  Each file defines a single chapter.  If the chapter is of the form `<filename>.<extension>`, for example `introduction.qmd`, the folders defined by `templateSearchPath` are searched, in order, and the first matching file is used.  This process allows a hierarchy of templates to be used.  If the chapter file is provided as an absolute or relative path, then the `templateSearchPath` is not used.  The file used is file in the precise location given.
* The `templateSearchPath` slot lists the folders that are searched, in order, to find the `chapter` files.
* The `variables` slot will be used to pass parameters to chapter files in a future version of `autoquarto`.  It is not currently used.
* The `type` slot defines the type of quarto object (eg `book` or `website`) that is being built.  The value is set, internally, by the object's constructor.

### Using the `templateSearchPath` hierarchy
Suppose a company provides global templates for components of reports at in a folder named `~/templates/gobal`.  Some of these global templates are named `index.qmd`,  `conclusion.qmd` and `environment.qmd`.  A project team has created project specific templates named `conclusion.qmd`, `main_body.qmd` and `intro.qmd` in a folder named `~/templates/projectA/`.  The project team is a running study named `study1` and has a provided study specific version of `intro.qmd` in `~/templates/projectA/study1`.  By default, the global template should be used.  However, if a project-level version of a template has been provided, then that should be used instead of the global version.  Similarly, a study-level template should override both a global or project-level template. 

This situation can be handled by setting the `templateSearchPath` slot to

```
list("~/templates/projectA/study1", "~/templates/projectA/", "~/templates/gobal")
```

If the `chapters` slot has the value

```
list(
  "index.qmd", 
  "intro.qmd", 
  "main_body.qmd", 
  "conclusion.qmd", 
  "environment.qmd", 
  "~/specialReports/appendix3.qmd"
)
```

Then

* `index.qmd` will be taken from `~/templates/gobal`
* `intro.qmd` will be taken from `~/templates/projectA/study1`
* `main_body.qmd` will be taken from `~/templates/projectA/`
* `conclusion.qmd` will be taken from `~/templates/projectA/`
* `environment.qmd` will be taken from `environment.qmd`
* `~/specialReports/appendix3.qmd` will be used as is

## Defining the `QuartoObject`
`autoquarto` defines constructors for the `QuartoBook` and `QuartoWebsite` classes.  The simplest use of the constructor is

```
myBook <- QuartoBook()
```

The `templateSearchPath` and `chapter` slots can be set when calling the constructor

```
myBook <- QuartoBook(
            templateSeachPath = list("~/templates/projectA/study1", "~/templates/projectA/", "~/templates/gobal"),
            chapters=list("index.qmd", "body.qmd", "conclusion.qmd")
          )
```

After initial creation of the object, these entries can be added to each slot with the `addTemplateSearchPath` and `addChapter` methods.

```
myBook <- addTemplateSearchPath(x, "~/templates/gobal")
myBook <- addChapter("environment.qmd")
```

All `QuartoObject` methods support piping.  The two statments above are equivalent to

```
library(magrittr)

myBook <- myBook %>% 
            addTemplateSearchPath(x, "~/templates/gobal") %>% 
            addChapter("environment.qmd")
```

By default, the `addXXXX`methods add new entries at the end of the current list, but the `.after` parameter can be used to insert the new entry at a different position in the list:

```
myBook <- myBook %>% addChapter("~/specialReports/appendix3.qmd", .after=2)
```

Existing elements of the list can be deleted with the appropriate `removeXXXX` method.  The following code returns `myBook` to its initial state:

```
myBook <- myBook %>% 
            removeTemplateSearchPath(x, "~/templates/gobal") %>% 
            removeChapter("environment.qmd") %>% 
            removeChapter("~/specialReports/appendix3.qmd")
```

## Publishing the `QuartoObject`
Once the methods discussed above have added the necessary information to the `QuartoObject`, it can be published by calling the `publish` method.

```
publish(myBook)
```

or 

```
publish(
  myBook,  
  outFile="/home/kirkpatj/Packages/autoquarto/tests/testOutput/testReport.html", 
)
```

or

```
publish(
  myBook, 
  outFile="/home/kirkpatj/Packages/autoquarto/tests/testOutput/testReport.html", 
  workDir="/home/kirkpatj/Packages/autoquarto/temp"
)
```

`publish`ing a `QuartoObject` also creates a log file (by default, in the same folder as `outFile`) describing what actions were taken by the `publish` method.  The log file has the same base name as the `outFile` with a timestamp appended.  For xample, in the examples above, the log file might be named `testReport_2023-02 Feb-10_143122.log`.

The detail and format of the messages in the log file can be controlled using standard calls to `futile.logger` functions, for example;

```
futile.logger::flog.threshold(futile.logger::DEBUG)
futile.logger::flog.layout(futile.logger::layout.format("~t ~l ~n ~f: ~m"))
futile.logger::flog.appender(futile.logger::appender.console()) 
```

and the content of a typical log file could be similar to

```
2023-02-10 14:36:08 INFO [autoquarto:publish]: Writing _quarto.yml
2023-02-10 14:36:15 DEBUG [autoquarto:publish]: _quarto.yml is:
2023-02-10 14:36:15 DEBUG [autoquarto:publish]: project:
2023-02-10 14:36:15 DEBUG [autoquarto:publish]:   type: book
2023-02-10 14:36:15 DEBUG [autoquarto:publish]:   output-dir: /home/kirkpatj/Packages/autoquarto/tests/testOutput
2023-02-10 14:36:15 DEBUG [autoquarto:publish]: book:
2023-02-10 14:36:15 DEBUG [autoquarto:publish]:   title: What is the title?
2023-02-10 14:36:15 DEBUG [autoquarto:publish]:   author: Who Is The Author
2023-02-10 14:36:15 DEBUG [autoquarto:publish]:   date: '`r format(Sys.Date())` at `r format(Sys.time())` on `r Sys.info()[['nodename']]`'
2023-02-10 14:36:15 DEBUG [autoquarto:publish]:   chapters:
2023-02-10 14:36:15 DEBUG [autoquarto:publish]:   - index.qmd
2023-02-10 14:36:15 DEBUG [autoquarto:publish]:   - intro.qmd
2023-02-10 14:36:15 DEBUG [autoquarto:publish]:   - body.qmd
2023-02-10 14:36:15 DEBUG [autoquarto:publish]:   - conclusion.qmd
2023-02-10 14:36:15 DEBUG [autoquarto:publish]:   - environment.qmd
2023-02-10 14:36:15 DEBUG [autoquarto:publish]:   output-file: testReport
2023-02-10 14:36:15 DEBUG [autoquarto:publish]: format:
2023-02-10 14:36:15 DEBUG [autoquarto:publish]:   html:
2023-02-10 14:36:15 DEBUG [autoquarto:publish]:     theme: cosmo
2023-02-10 14:36:15 DEBUG [autoquarto:publish]:   pdf: default
2023-02-10 14:36:15 DEBUG [autoquarto:publish]: 
2023-02-10 14:36:15 INFO [autoquarto:publish]: Copying chapter files...
2023-02-10 14:36:15 DEBUG [autoquarto:publish]: Copying `/home/kirkpatj/Packages/autoquarto/tests/testData/global/index.qmd` to `/home/kirkpatj/Packages/autoquarto/temp`...
2023-02-10 14:36:15 DEBUG [autoquarto:publish]: Copying `/home/kirkpatj/Packages/autoquarto/tests/testData/projectA/intro.qmd` to `/home/kirkpatj/Packages/autoquarto/temp`...
2023-02-10 14:36:15 DEBUG [autoquarto:publish]: Copying `/home/kirkpatj/Packages/autoquarto/tests/testData/global/body.qmd` to `/home/kirkpatj/Packages/autoquarto/temp`...
2023-02-10 14:36:15 DEBUG [autoquarto:publish]: Copying `/home/kirkpatj/Packages/autoquarto/tests/testData/projectA/study1/conclusion.qmd` to `/home/kirkpatj/Packages/autoquarto/temp`...
2023-02-10 14:36:15 DEBUG [autoquarto:publish]: Copying `/home/kirkpatj/Packages/autoquarto/tests/testData/global/environment.qmd` to `/home/kirkpatj/Packages/autoquarto/temp`...
2023-02-10 14:36:15 INFO [autoquarto:publish]: Rendering report...
2023-02-10 14:36:31 INFO [autoquarto:publish]: Done
2023-02-10 14:36:31 INFO [autoquarto:publish]: Exit - QuartoObject

```


## Notes
* Your report must include an `index.qmd` file.  This is a limitation of Quarto.  See [here](https://github.com/quarto-dev/quarto-cli/discussions/4024) for more information.
* `output-file` affects only single-file destinations such as PDF.  It is ignored for multi-file destinations such as html.  For multi-file destinations, the project main file is always (eg) index.html.

## To do
* Support `param`eters
* Provide support for Parts
* Implement support for bibliographies
* Implement support for book- and website-specific features
* Improve test coverage
* Provide getters (and setters) for all class clots
